---
title: "Data 607 Ass2"
author: "Kai Lukowiak"
date: '2017-10-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NYC Crashes:

Import the data:

```{r}
library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)
library(ggthemes)
```

```{r}
nycDF <-  read_excel("/Users/kailukowiak/OneDrive - CUNY School of Professional Studies/DATA 607 Repository/Project 2/cityacc-en-us.xlsx")
head(nycDF)
tail(nycDF)

```

```{r}
nycDF <- nycDF %>% filter(!is.na(X__1))
head(nycDF)
tail(nycDF)
```

Now we need to add row names.

```{r}

colnames(nycDF) <- nycDF[1,]
nycDF <- nycDF[-1,]
head(nycDF) 

```


Everthing is reletively pretty now, but we have citywide, borough, and precicnt data. This isn’t very tidy.

Seperate borough and precent data.

```{r}
borDF <- nycDF %>% 
  filter( !grepl("\\d", GeoCode))
preDF <- nycDF %>% 
  filter( grepl("\\d", GeoCode))
```

Create new column.

```{r}
nycDF <- nycDF %>% mutate(Borough = GeoCode)
```

Remove numerics from Borough

```{r}
nycDF$Borough <- sub("\\d\\d\\d", NA , nycDF$Borough)
nycDF <- nycDF %>% fill(Borough, .direction = "down")# Add factor variables. 
precDF <- nycDF %>% 
  filter( grepl("\\d", GeoCode))
precDF %>%
  head() 
```

```{r}
precDF$Borough <-  precDF$Borough %>% recode(M = "Manhatten",
         B = "Bronx",
         K = "Brooklyn",
         Q = "Queens",
         S = "StattenIsland") # Adds names.
precDF$Borough <- as.factor(precDF$Borough)
cols <- colnames(precDF)
precDF[, cols[-c(2,15)]] <-  as.numeric(unlist(precDF[, cols[-c(2,15)]])) # Convert to numeric
```


# Analysis of NYC Crashes

```{r}
plt <-  precDF %>% 
  group_by(Borough) %>% 
  summarise(numAccidents = sum(Number_of_Motor_Vehicle_Collisions))

ggplot(plt, aes(x = Borough, y = numAccidents))+
  geom_bar(stat = 'identity') + 
  xlab("Borough") +
  ylab("Number of Accidents") +
  ggtitle("Number of Accidents by Borough") +
  theme_fivethirtyeight()
```

Special thanks to [Rob Barry](http://rob-barry.com/2015/06/14/Mapping-in-R/) for nyc precincts.



```{r}
# Grab the shapefile and extract it
# You may want to add some file paths here
download.file(
  # I saved a version on my server so this script should work
  # even if the NYC planning department changes things around
  "http://www.rob-barry.com/assets/data/mapping/nypp_15b.zip",
  destfile = "nypp_15b.zip"
)
unzip(zipfile = "nypp_15b.zip")

# Now, load package to read the shapefile
library("rgdal")

# Read it into an sp object:
nypp <- readOGR("nypp_15b", "nypp")
```


```{r}
library("RColorBrewer")
library("classInt")

# Plot it
pal <- brewer.pal(5, "YlOrRd")
fill.clr <-
  findColours(
    classIntervals(as.numeric(precDF$Number_of_Motor_Vehicle_Collisions), style = "pretty", n = 5),
    pal
  )
plot(nypp, col = fill.clr)
legend(
  "topleft",
  fill=attr(fill.clr, "palette"),
  legend=names(attr(fill.clr, "table")),
  bty = "n"
)
title("Accidents by Precinct")
```


```{r}
# Plot it
pal <- brewer.pal(5, "YlOrRd")
fill.clr <-
  findColours(
    classIntervals(as.numeric(precDF$Injury_or_Fatal_Collisions), style = "pretty", n = 5),
    pal
  )
plot(nypp, col = fill.clr)
legend(
  "topleft",
  fill=attr(fill.clr, "palette"),
  legend=names(attr(fill.clr, "table")),
  bty = "n"
)
title("Total Number of Casualties")

```

# Hospital Data

Importing the data:

```{r}
hosDF <- read.csv("/Users/kailukowiak/OneDrive - CUNY School of Professional Studies/DATA 607 Repository/Project 2/Complications and Deaths - Hospital.csv")

```

There are a lot of NA values. First we need to change them to NA
```{r}
hosDF[ hosDF == "Not Available" ] <- NA
```


We can remove NA’s but this gives us an incomplete dataset. It might be better to remove the entire hospital if it is missing one value. We will deal with this later.

Right now we need to find the length of time that the hospitals studied

```{r}
osDF <- hosDF %>% mutate(timeDiff = difftime(as.POSIXct(as.Date(hosDF$Measure.End.Date,
                                               format = '%m/%d/%Y')), 
         as.POSIXct(as.Date(hosDF$Measure.Start.Date, 
                                               format = '%m/%d/%Y')), 
         units = 'days'))
```

```{r}
hosDF$Score <-  as.numeric(hosDF$Score)
plt <- hosDF %>% 
  group_by(Measure.Name) %>% 
  summarise(facSum = sum(Score, na.rm = T))
 
ggplot(plt, aes(Measure.Name, facSum))+
  geom_bar(stat = 'identity') +
  theme(axis.text.x=element_text(angle=90, hjust=1))+ coord_flip()+
  ggtitle('Deaths by Cause')
```

Create dummy variables for Measure ID.

```{r}
hosDF %>% 
  mutate(yesno = 1) %>% 
  distinct() %>% 
  spread(Measure.ID, yesno, fill = 0) %>% 
  glimpse()
```

# Chronic Disease

```{r}
library(readr)
U_S_Chronic_Disease_Indicators_CDI_ <- read_csv("/Users/kailukowiak/U.S._Chronic_Disease_Indicators__CDI_ (1).csv")
```

```{r}
chronDF <- U_S_Chronic_Disease_Indicators_CDI_

chronDF <- chronDF %>% select_if(~sum(!is.na(.)) > 0) # Removes all rows that only have NA values
chronDF$DataValueFootnoteSymbol <- NULL
chronDF$DatavalueFootnote <- NULL
chronDF$DataValueUnit <- NULL # Removing not very useful data
chronDF <- filter(chronDF, LocationDesc != c("Guam", "District of Columbia","Puerto Rico",
                                             "United States","Virgin Islands"))

library(fiftystater)
library(mapproj)
```





```{r}
alcohol <- filter(chronDF, Question == "Alcohol use among youth" & 
                    YearStart == 2015 & StratificationCategory1 == "Overall")
alcohol$LocationDesc <- tolower(alcohol$LocationDesc) # Needs to be lowercase to match the map.
ggplot(alcohol, aes(map_id = LocationDesc)) +
  geom_map(aes(fill = DataValue), map = fifty_states) +
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())+
  ggtitle("Alcohol use among youth")
```

```{r}
alcoholFem <- filter(chronDF, Question == "Heavy drinking among women aged 18-44 years" & 
                       YearStart == 2015 & StratificationCategory1 == "Overall")
alcoholFem$LocationDesc <- tolower(alcoholFem$LocationDesc) # Needs to be lowercase to match the map.
ggplot(alcoholFem, aes(map_id = LocationDesc)) +
  geom_map(aes(fill = DataValue), map = fifty_states) +
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank()) +
  ggtitle("Heavy drinking among women aged 18-44 years")
```

Nevada doesn’t make a lot of sense being so low, but maybe my sterotypes of Americans need updating. It’s interesting that there seems to be little correlation between youth alcohol consumption and heavy drinking in adulthood.