---
title: "Week 3 Assigment Data 607"
author: "Kai Lukowiak"
date: '2017-09-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assigment 3
## Q 3
Copy the introductory example. The vector `name` stores the extracted names.
```{r}
library(stringr)
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name # This was taken from the textbook.
```

### a)
Use the tools of this chapter to rearrange the vector so that all elements conform to the
standard first_name last_name.

I am assuming that the instructions demand that honerifics are also removed. (I.e. J.K. Rowling stays the same but Ms. J.K. Rowling drops the Ms.)
```{r}
name

```

```{r}
name2 <- str_replace_all(name, "(.+)(, .+)$", "\\2 \\1") # Change order
name2 
name3 <- str_replace_all(name2, ", ", "") # Remove commas
name3
name4 <- str_replace_all(name3, "[A-Z][a-z]([a-z]?)\\.", "") # Remove honerifics.
name4
```

### b) 
Construct a logical vector indicating whether a character has a title (i.e., Rev. and
Dr.).

Well I regret not reading this question before putting all that effor into removing them in the first place. 

```{r}
library(knitr)
df <- data.frame(name3)
df$title <- str_detect(string = name3, pattern = "\\w{2,3}\\.")
kable(df)
```
### c) 
Construct a logical vector indicating whether a character has a second name.

```{r}
df$secname <- str_detect(string = name3, pattern = "[A-Z]{1}\\.")
kable(df)
```

It should be noted that this method is not very robust. If the data was more complex an additional column would be made to count the number of spaces without a proceeding period (to ignore the honerifics) and then the new column and the one I just created would be added together and if there was one true, we would know that the character had a middle name.

## Q4
Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

### (a) `[0-9]+\\$`
```{r}
test <- c("gjkaef123892389$fsafsdlkj", "asfdlk$afsdlk", "234$123", "1234", "$23sf")
test1 <- unlist(str_extract_all(test, pattern = "[0-9]+\\$" ))
test1
```
From these examples we can see that the expression only brings a number and a $. And only in that speficic order.

### (b) `\\b[a-z]{1,4}\\b`
This will detect a word beging with 1-4 characters and another word edge. In other words any word that is between 1 and four chacters. Characters must be lowercase:
```{r}
test <- c("abcdefg", "abe def hgi", "abdefkl asl 234", "1234", "WTF", "abc")
test1 <- str_extract_all(test, pattern = "\\b[a-z]{1,4}\\b" )
test1
```

### (c) `.*?\\.txt$`


### (d) `\\d{2}/\\d{2}/\\d{4}`
### (e) `<(.+?)>.+?</\\1>`