---
title: "DATA_607_Ass1"
author: "Kai Lukowiak"
date: '2017-08-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Assigment 1
Very often, we’re tasked with taking data in one form and transforming it for easier downstream analysis. We will
spend several weeks in this course on tidying and transformation operations. Some of this work could be done in
SQL or R (or Python or…). Here, you are asked to use R—you may use base functions or packages as you like.
Mushrooms Dataset. A famous—if slightly moldy—dataset about mushrooms can be found in the UCI repository
here: https://archive.ics.uci.edu/ml/datasets/Mushroom. The fact that this is such a well-known dataset in the
data science community makes it a good dataset to use for comparative benchmarking. For example, if someone
was working to build a better decision tree algorithm (or other predictive classifier) to analyze categorical data,
this dataset could be useful. A typical problem (which is beyond the scope of this assignment!) is to answer the
question, “Which other attribute or attributes are the best predictors of whether a particular mushroom is
poisonous or edible?”
Your task is to study the dataset and the associated description of the data (i.e. “data dictionary”). You may need
to look around a bit, but it’s there! You should take the data, and create a data frame with a subset of the columns
in the dataset. You should include the column that indicates edible or poisonous and three or four other columns.
You should also add meaningful column names and replace the abbreviations used in the data—for example, in the
appropriate column, “e” might become “edible.” Your deliverable is the R code to perform these transformation
tasks.

## Loading some packages:

```{r}
library(knitr)
```

## Loading the data:
```{r}
df <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")
#Converting to a df.
df <- as.data.frame(df)
kable(head(df))

```

This doesn't look like much. Let's move on the naming the variables/features. 

## Meaningful Names:
The homework gave various tasks in a certain order but I think it makes sense to at least assign names to colums first. 

In my quick search I found this [RPub](https://rpubs.com/dvillalobos/DATA607-hmwk1) which figgured out a concise way to pull the data from the dictionary and name the columns. It also named the coded values with their proper name. The value added work that I am bringing to this assigment is that will automate naming the catagorical variables so that all information will be correctly encoded. 

```{r}
file <- 'https://raw.githubusercontent.com/dvillalobos/MSDA/master/607/Homework/Villalobos-Homework1-dictionary.txt'
mushroomsdict <- read.table(file, sep="|", header=TRUE, stringsAsFactors = FALSE)
mushroomsdict

```
This allows us to name the columns with the atribute column:
```{r}
colnames(df) <- mushroomsdict$Attribute
head(df)
```


## Naming the categories:
Again thanks to [Duubar Villalobos Jimenez](https://rpubs.com/dvillalobos/DATA607-hmwk1)
```{r}
transMush <- function(headcols){
  # Reading information: 
  mushHeadVals <- mushroomsdict$Information[headcols]
  # Must use the as.character to split into string.
  mushHeadVals <- strsplit(as.character(mushHeadVals), ",", fixed = TRUE) 
  # Convert to a data frame:
  mushHeadVals <- data.frame(mushHeadVals)
  # Make sure that the names carry over so that they can be matched with the values in df:
  colnames(mushHeadVals) <- mushroomsdict$Attribute[headcols]
  # seperate the values based on the "="
  mushHeadVals <- data.frame(do.call("rbind", 
                                     strsplit(as.character(mushHeadVals[,1]), 
                                              "=", fixed = TRUE)))
  # This command remanes the values:
  colnames(mushHeadVals) <- c(mushroomsdict$Attribute[headcols], "values")
  #assigninf the factor data type in a new data frame:
  mush[, headcols] <- factor(mush[, headcols], ordered = TRUE)
  levels(mush[, headcols]) <- as.character(mushHeadVals[,1])
  return(mush)
}
#New data set is, as far as I can tell not necessary but in order to preserve the integrity of the original it seems worth while. 
mush <- subset(df, select = c(1:dim(df)[2]))
head(mush)
```
## Value added:
In order to distiguish this analysis, I'm going to use a for loop to itterate over all coded values to see the true name. (It would be nice to do this with a vector applied to the function but I'm not good enough yet.)
```{r}
for(i in 1:23){
  mush <- transMush(i)
}
kable(head(mush, 20))
```


## Subsetting:
At this point, subsetting is easy. 
```{r}
names(mush)
subMush <- subset(mush, select = c("class", "cap-shape", "habitat", "ring-number"))
```
```{r}
kable(head(subMush, 20))
```

