---
title: "Project1"
author: "Kai Lukowiak"
date: '2017-09-20'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Project 1

The project can be read from the PDF posted on [github.com](https://github.com/kaiserxc/DATA_607/tree/master/Project%201)


## Loading libraries

The necessary libraries. 
```{r}
library(stringr)
library(knitr)
library(dplyr)
```

First we have to load the data. I'm going to load it from github.com so that the project will run regardless of the comptuer directory.

## Loading data

```{r}
theUrl <- "https://raw.githubusercontent.com/kaiserxc/DATA_607/master/Project%201/tournamentinfo.txt"
txtFile <- read.csv(theUrl, header = FALSE)
txtFile[c(1:10),] # This command was used because head() did not behave.

tail(txtFile) # Looking at the tail function it appears to work in the MD
# file but not here. I'll leave both in to show diff methods.

```

## Cleaning data
The girst 4 rows contain info on the turnament and are superflous to our project. 


```{r}
txtFile1 <-txtFile[-c(1:4),]
txtFile1 %>% head() 
class(txtFile1)

```
We need to change this to character

```{r}
txtFile1 <- as.character(txtFile1)
class(txtFile1)
```



Success!


We need the first and second rows out of every three (discard the third). We also need to seperate them. 
```{r}
name <- txtFile1[seq(from = 1, to = length(txtFile1), by = 3)]
rating <- txtFile1[seq(from = 2, to = length(txtFile1), by = 3)]

```


## Extracting relevant data

```{r}
p_id <- as.integer(str_extract(name, "\\d+")) # could have just done c(1:64)
p_name <- str_trim(str_extract(name, "(\\w+\\s){2,3}")) #This works because name is the first instance.
p_point <- as.numeric(str_extract(name, "\\d.\\d"))
p_rate <- as.integer(str_extract(str_extract(rating, "\\D\\d{3,4}\\D"), "\\d+")) 
state <- str_extract(rating, "\\w\\w")
opp_id <- str_extract_all(str_extract_all(name, "\\d+\\|"), "\\d+")

```

## Turning the list into a data frame 

Thanks to [this](found herehttps://stackoverflow.com/questions/15201305/how-to-convert-a-list-consisting-of-vector-of-different-lengths-to-a-usable-data)
```{r}
n.obs <- sapply(opp_id, length)
seq.max <- seq_len(max(n.obs))
mat <- t(sapply(opp_id, "[", i = seq.max)) 
df2 <- as.data.frame(mat)
```

## Replacing the id's with ratings

Special thanks to [this stack overflow](https://stackoverflow.com/questions/35636315/replace-values-in-a-dataframe-based-on-lookup-table) for helping me. 
```{r}
key <- data.frame(p_id, p_rate)
df3 <- df2
df3[] <- key$p_rate[match(unlist(df2), key$p_id)]
kable(head(df3))
```


## Calculating the average

```{r}
df3$oppAvg <- round(apply(df3, 1, mean, na.rm=TRUE))
kable(head(df3))
```


```{r}
df <- data_frame(p_id, p_name, state, p_point, p_rate, df3$oppAvg )
colnames(df) <-  c("ID", "Player_Name", "State", "Point", "Pre_Match_Rating", "Opponent_Average")
kable(head(df, 10))
```
## Write the CSV

```{r}
write.csv(df, file = "/Users/kailukowiak/OneDrive - CUNY School of Professional Studies/DATA 607 Repository/Project 1/clean_tournamentinfo.txt")
```


## Does chess rank people

Let's do a regression to find out.

```{r}
fit <- lm(Opponent_Average ~ Pre_Match_Rating, data = df)
summary(fit)
```

```{r}
cor(x = df$Pre_Match_Rating, y = df$Opponent_Average)
```

This might suggest otherwise.

```{r}
fit <- lm(Point ~ Pre_Match_Rating, data = df)
summary(fit)
```

No surprise here. Being a good player means you score more.


