---
title: "Webscraping to Find Key DS Skills"
author: "Kai Lukowiak"
date: '2017-10-28'
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data 607 Week 9 Assignment
## webscraping from an API
### Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.

* Data can be collected [here](http://developer.nytimes.com/docs)


```{r warning=FALSE, error=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(jsonlite)
library(httr)
library(pander)
```

The API key and url to extract the most popular opinion peices over the last thirty days are here. 

```{r}
apiKey <-  "&api-key=b95585fd6b264ab08e28fb0e96998aae"
urlBase <-"https://api.nytimes.com/svc/mostpopular/v2/mostshared/Opinion/30.json?"
nytURL <- paste(urlBase, apiKey, sep = "")
```

We use the GET command to extract a json like object from the API

```{r}
nytData <- GET(nytURL)
content(nytData, 'text')

```

And convert it into a data frame with relevant columns.

```{r}
json_raw <- content(nytData, type = "text", encoding = "UTF-8")
json1 <- fromJSON(json_raw)
df <- as.data.frame(json1$results)
df1 <- df %>% select(byline, title, abstract, published_date)
df1$byline <-  str_replace(df1$byline, 'By ', "")
df1 %>% kable("html") 
  
  
  
```


